Naturalness in U(1) extensions of the MSSM

### 2014-12-03 paper/e6ssm-tuning-paper.tex
Added more components of the tree level master formula to the appendix, added an extra pair of brackets and in-line equation, removed extra bracket in sentence, changed m_z --> M_Z for consistency with previous tuning paper

The Higgs causes a naturalness problem because at tree level it has an
upper bound of the mass of the $Z$ boson, $m_Z$. The dominant higher
order corrections to the Higgs mass come from stops and to obtain a
$125$ GeV Higgs they need be rather heavy. Heavy stops will provide a
large contribution to the soft breaking mass for the up-type Higgs
scalar, $m_{H_u}^2$ at low energies, through the evolution of the
renormalisation group equations (RGEs) from the grand unification
(GUT) scale to the electroweak (EW) scale. This affects the SUSY
prediction of the electroweak VEV, v, or $m_Z$.  This naturalness
problem motivates both further examination of non-minimal SUSY models
that can raise the Higgs mass without the need for heavy stops and
alternative possibilities for how the soft breaking parameters get
generated, which might set them them at lower energies, reducing the
influence the stops have on the soft breaking mass for the up-type
Higgs scalar, $m_{H_u}^2$.

@@

The breaking of $E_6$ into $SO(10)$, gives rise to $E_6\to
SO(10)\times U(1)_{\psi}$, and the subsequent breaking of $SO(10)$
into $SU(5)$, gives $SO(10)\to SU(5)\times U(1)_{\chi}$ (this is
reviewed in e.g.~\cite{Langacker:2008yv})). The extra $U(1)$ gauge symmetry at low energies should then be a linear combination of these in the $E_6$ inspired case,
\be
U(1)^\prime = \cos(\theta) U(1)_{\chi} + \sin(\theta) U(1)_{\psi}.
\ee In table \ref{tab:E6charges} the charges for several popular $E_6$ inspired $U(1)$ extensions are shown.
% Table of U(1)' charges, which may or may not be useful to include.
\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cccccccccccccc}
 & $\hat{Q}$ & $\hat{u}^C$ & $\hat{d}^C$ & $\hat{L}$ & $\hat{e}^C$ & $\hat{N}^C$ & $\hat{S}$ & $\hat{H}_2$ & $\hat{H}_1$ & $\hat{D}$ & $\hat{\overline{D}}$ & $\hat{H}'$ & $\hat{\overline{H'}}$ \\[1mm]
\hline
$\sqrt{\frac{5}{3}}Q_i^Y$ & $\frac{1}{6}$ & $-\frac{2}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $1$ & $0$ & $0$ & $\frac{1}{2}$ & $-\frac{1}{2}$ & $-\frac{1}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $\frac{1}{2}$ \\[1mm]
%\hline
$2\sqrt{6}Q_i^\psi$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $4$ & $-2$ & $-2$ & $-2$ & $-2$ & $1$ & $-1$\\[1mm]
%\hline
$2\sqrt{10}Q_i^\chi$ & $-1$ & $-1$ & $3$ & $3$ & $-1$ & $-5$ & $0$ & $2$ & $-2$ & $2$ & $-2$ & $3$ & $-3$\\[1mm]
%\hline
$\sqrt{40}Q_i^N$ & $1$ & $1$ & $2$ & $2$ & $1$ & $0$ & $5$ & $-2$ & $-3$ & $-2$ & $-3$ & $2$ & $-2$ \\[1mm]
\end{tabular}
\end{ruledtabular}
\caption{The $U(1)_Y$, $U(1)_\psi$, $U(1)_\chi$ and $U(1)_N$ charges of the chiral superfields in the $E_6$ model. The specific case of $U(1)_N$, corresponding to the E$_6$SSM, is obtained for $\theta_{E_6}=\arctan\sqrt{15}$.}
\label{tab:E6charges}
\end{table}

@@

Nonetheless while this model is very elegant so far, the  superpotential, \ref{Eq:e6-superpot} contains dangerous terms which can induce proton decay and lead to large flavour changing neutral currents (FCNCs).  There are a number of approaches to suppress these terms, involving the use of different discrete symmetries.  Here for the purposes of renormalisation group running we will simply include the following unsuppressed superpotential terms, which follows the approach taking in work on the cE$_6$SSM \cite{Athron:2009ue, Athron:2009bs},
%
%% \begin{align}
%%   \begin{split}
    \ba
    W &\approx& - y_{\tau} (\SuperField{H}_d
    \SuperField{L}_3) \SuperField{e}^C_3 - y_b (\SuperField{H}_d
    \SuperField{Q}_3) \SuperField{d}_3^C - y_t (\SuperField{Q}_3
    \SuperField{H}_u) \SuperField{u}_3^C\\
    &&
    + \lambda_i \SuperField{S} (\SuperField{H}_{1i}
    \SuperField{H}_{2i}) + \kappa_i \SuperField{S} \SuperField{D}_i
    \SuperField{\overline{D}}_i + \mu' (\SuperField{H}'
    \SuperField{\overline{H'}}),
    \ea
  %% \end{split}
  \label{SuPot_RGE}
%%\end{align}
%

@@

with $f_1,f_2,f_3$ as given in Eq. (\ref{eq:E6EWSBConditions}). For a running parameter $q$ appearing in the tree level EWSB conditions, the corresponding contribution to the sensitivity coefficient may be written 

@@

C=\frac{1}{2}\Big ( z_1\frac{\partial f_1}{\partial M_Z}+z_2\frac{\partial f_2}{\partial M_Z}+z_3\frac{\partial f_3}{\partial M_Z}\Big ).

@@

@@

Provided that one does not run over the too large a range of scales, the solutions to the

@@

b_q^{(2)}(M_X)=\frac{1}{2!}\left [ \sum_{q_k}\beta_{q_k}^{(1)}\frac{\partial \beta_q}{\partial q_k}\right ]_{M_X}

@@

and the notation $[\ldots]_Q$ indicates that everything within the brackets is evaluated at the scale $Q$. It is straightforward 
to get the explicit series expansions for the MSSM parameters listed above. Due to the smallness of the first and second

>>>>

The Higgs causes a naturalness problem because at tree level it has an
upper bound of the mass of the $Z$ boson, $M_Z$. The dominant higher
order corrections to the Higgs mass come from stops and to obtain a
$125$ GeV Higgs they need be rather heavy. Heavy stops will provide a
large contribution to the soft breaking mass for the up-type Higgs
scalar, $m_{H_u}^2$ at low energies, through the evolution of the
renormalisation group equations (RGEs) from the grand unification
(GUT) scale to the electroweak (EW) scale. This affects the SUSY
prediction of the electroweak VEV, $v$, or $M_Z$.  This naturalness
problem motivates both further examination of non-minimal SUSY models
that can raise the Higgs mass without the need for heavy stops and
alternative possibilities for how the soft breaking parameters get
generated, which might set them them at lower energies, reducing the
influence the stops have on the soft breaking mass for the up-type
Higgs scalar, $m_{H_u}^2$.

@@

The breaking of $E_6$ into $SO(10)$, gives rise to $E_6\to
SO(10)\times U(1)_{\psi}$, and the subsequent breaking of $SO(10)$
into $SU(5)$, gives $SO(10)\to SU(5)\times U(1)_{\chi}$ (this is
reviewed in e.g.~\cite{Langacker:2008yv}). The extra $U(1)$ gauge symmetry at low energies should then be a linear combination of these in the $E_6$ inspired case,
\be
U(1)^\prime = \cos(\theta) U(1)_{\chi} + \sin(\theta) U(1)_{\psi}.
\ee In table \ref{tab:E6charges} the charges for several popular $E_6$ inspired $U(1)$ extensions are shown.
% Table of U(1)' charges, which may or may not be useful to include.
\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cccccccccccccc}
 & $\hat{Q}$ & $\hat{u}^C$ & $\hat{d}^C$ & $\hat{L}$ & $\hat{e}^C$ & $\hat{N}^C$ & $\hat{S}$ & $\hat{H}_2$ & $\hat{H}_1$ & $\hat{D}$ & $\hat{\overline{D}}$ & $\hat{H}'$ & $\hat{\overline{H'}}$ \\[1mm]
\hline
$\sqrt{\frac{5}{3}}Q_i^Y$ & $\frac{1}{6}$ & $-\frac{2}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $1$ & $0$ & $0$ & $\frac{1}{2}$ & $-\frac{1}{2}$ & $-\frac{1}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $\frac{1}{2}$ \\[1mm]
%\hline
$2\sqrt{6}Q_i^\psi$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $4$ & $-2$ & $-2$ & $-2$ & $-2$ & $1$ & $-1$\\[1mm]
%\hline
$2\sqrt{10}Q_i^\chi$ & $-1$ & $-1$ & $3$ & $3$ & $-1$ & $-5$ & $0$ & $2$ & $-2$ & $2$ & $-2$ & $3$ & $-3$\\[1mm]
%\hline
$\sqrt{40}Q_i^N$ & $1$ & $1$ & $2$ & $2$ & $1$ & $0$ & $5$ & $-2$ & $-3$ & $-2$ & $-3$ & $2$ & $-2$ \\[1mm]
\end{tabular}
\end{ruledtabular}
\caption{The $U(1)_Y$, $U(1)_\psi$, $U(1)_\chi$ and $U(1)_N$ charges of the chiral superfields in the $E_6$ model. The specific case of $U(1)_N$, corresponding to the E$_6$SSM, is obtained for $\theta=\arctan\sqrt{15}$.}
\label{tab:E6charges}
\end{table}

@@

Nonetheless while this model is very elegant so far, the  superpotential, (\ref{Eq:e6-superpot}) contains dangerous terms which can induce proton decay and lead to large flavour changing neutral currents (FCNCs).  There are a number of approaches to suppress these terms, involving the use of different discrete symmetries.  Here for the purposes of renormalisation group running we will simply include the following unsuppressed superpotential terms, which follows the approach taking in work on the cE$_6$SSM \cite{Athron:2009ue, Athron:2009bs},
%
%% \begin{align}
%%   \begin{split}
    \ba
    W &\approx& - y_{\tau} (\SuperField{H}_d
    \SuperField{L}_3) \SuperField{e}^C_3 - y_b (\SuperField{H}_d
    \SuperField{Q}_3) \SuperField{d}_3^C - y_t (\SuperField{Q}_3
    \SuperField{H}_u) \SuperField{u}_3^C\\
    &&
    + \lambda_i \SuperField{S} (\SuperField{H}_{1i}
    \SuperField{H}_{2i}) + \kappa_i \SuperField{S} \SuperField{D}_i
    \SuperField{\overline{D}}_i + \mu' (\SuperField{H}'
    \SuperField{\overline{H'}}),
    \ea
  %% \end{split}
  \label{SuPot_RGE}
%%\end{align}
%

@@

with $f_1,f_2,f_3$ as given in Eq. (\ref{eq:E6EWSBConditions}). Explicitly, we find that
\begin{eqnarray*}
z_1&=&%\frac{M_Z\cos^2\beta}{2\bar{g}^3}\Big [ 4M_Z^2\Big ( \sqrt{2}\lambda A_\lambda \cos\beta-2s(\lambda^2+g_1'^2\tilde{Q}_2\tild%e{Q}_S )\sin\beta\Big )\nonumber\\
%&& \times \Big ( \sqrt{2}\lambda A_\lambda \cos 2\beta+g_1'^2(\tilde{Q}_1-\tilde{Q}_2)\tilde{Q}_Ss\sin 2\beta \Big )\Big ]
\end{eqnarray*}
For a running parameter $q$ appearing in the tree level EWSB conditions, the corresponding contribution to the sensitivity coefficient may be written 

@@

C=\frac{1}{2}\left ( z_1\frac{\partial f_1}{\partial M_Z}+z_2\frac{\partial f_2}{\partial M_Z}+z_3\frac{\partial f_3}{\partial M_Z}\right ),

@@

with 
\ba
\frac{\partial f_1}{\partial M_Z}&=&\frac{2}{\bar{g}}\,\cos\beta\left ( m_{H_d}^2+\frac{\lambda^2s^2}{2}+\frac{g_1'^2}{2}\tilde{Q}_1\tilde{Q}_Ss^2+\frac{6g_1'^2}{\bar{g}^2}\tilde{Q}_1^2M_Z^2\right )-\sqrt{2}\frac{\lambda A_\lambda s}{\bar{g}}\,\sin\beta\nonumber \\
&&+\frac{3M_Z^2}{\bar{g}}\,\cos\beta\,\cos 2\beta+\frac{6}{\bar{g}^3}M_Z^2\,\sin\beta\,\sin 2\beta\left [ \lambda^2+g_1'^2\left(\tilde{Q}_1\tilde{Q}_2-\tilde{Q}_1^2\right )\right ],\nonumber\\
\frac{\partial f_2}{\partial M_Z}&=&\frac{2}{\bar{g}}\,\sin\beta\left ( m_{H_u^2}+\frac{\lambda^2s^2}{2}+\frac{g_1'^2}{2}\tilde{Q}_2\tilde{Q}_Ss^2+\frac{6g_1'^2}{\bar{g}^2}\tilde{Q}_2^2M_Z^2\right )-\sqrt{2}\frac{\lambda A_\lambda s}{\bar{g}}\,\cos\beta \nonumber\\
&&-\frac{3M_Z^2}{\bar{g}}\,\sin\beta\,\cos 2\beta+\frac{6}{\bar{g}^3}M_Z^2\,\cos\beta\,\sin 2\beta\left [ \lambda^2+g_1'^2\left ( \tilde{Q}_1\tilde{Q}_2-\tilde{Q_2}^2\right )\right ],\nonumber \\
\frac{\partial f_3}{\partial M_Z}&=&\frac{4M_Z}{\bar{g}^2}\left [ \lambda^2s-\frac{\lambda A_\lambda}{\sqrt{2}}\,\sin 2\beta+g_1'^2\tilde{Q}_Ss\left ( \tilde{Q}_1\,\cos^2\beta+\tilde{Q}_2\,\sin^2\beta \right )\right ].\nonumber
\ea

@@

Provided that one does not run over too large a range of scales, the solutions to the

@@

b_q^{(2)}(M_X)=\frac{1}{2!}\left . \sum_{q_k}\beta_{q_k}^{(1)}\frac{\partial \beta_q}{\partial q_k}\right |_{M_X}

@@

It is straightforward to get the explicit series expansions for the MSSM parameters listed above. Due to the smallness of the first and second
